---
title: "By-Locus Summaries"
author: "Eric Archer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Summarizing Data in strataG"}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
There are several by-locus summary functions available for _gtypes_ objects. Given some sample microsatellite data:
```{r}
options(digits = 2)
library(strataG.devel)
data(dolph.strata)
data(dolph.msats)
msats.merge <- merge(dolph.strata, dolph.msats, all.y = TRUE, description = date())
msats <- df2gtypes(msats.merge, ploidy = 2, id.col = 1, strata.col = 3, loc.col = 5)
rownames(dolph.strata) <- dolph.strata$ids
schemes(msats) <- dolph.strata[, c("fine", "broad")]
msats <- stratify(msats, "broad")
msats <- subset(msats, loci = locNames(msats)[1:4])
```

One can calculate the following summaries:

The number of alleles at each locus:
```{r}
numAlleles(msats)
```

The number of samples with missing data at each locus:
```{r}
numMissing(msats)
```
which can also be expressed as a proportion of samples with missing data:
```{r}
numMissing(msats, prop = TRUE)
```

The allelic richness, or the average number of alleles per sample:
```{r}
allelicRichness(msats)
```

The observed and expected heterozygosity:
```{r}
obsvdHet(msats)
exptdHet(msats)
```

The percent of alleles that are unique (present in only one sample):
```{r}
pctUniqueAlleles(msats)
```

The value of theta based on heterozygosity:
```{r}
theta(msats)
```

These measures are all calculated in the _summarizeLoci_ function and returned as a matrix. This function also allows you to calculate the measures for each stratum separately, which returns a list for each stratum:
```{r}
summarizeLoci(msats)
summarizeLoci(msats, by.strata = TRUE)
```

One can also obtain the allelic frequencies for each locus overall and by-strata by:
```{r}
alleleFreqs(msats)
alleleFreqs(msats, by.strata = TRUE)
```
