---
title: "QA/QC"
author: "Eric Archer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{5. QA/QC}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
```{r echo = FALSE, message = FALSE}
options(digits = 2)
library(strataG)
```
Quality of genotypes and sequences can be assessed with several functions. For genotypic data, by-locus summaries can be conducted with the _summarizeLoci_ function, which will also produce summaries for each strata:
```{r}
data(msats.g)
msats <- msats.g
smry <- summarizeLoci(msats)
head(smry)
```


The _dupGenotypes_ function identifies samples that have the same or nearly the same genotypes. The number (or percent) of loci that must be shared in order for it to be considered a duplicate can be set by the _num.shared_ argument. The return data.frame provides which loci the two samples show mismatches at so they can be reviewed.
```{r}
# Find samples that share alleles at 2/3rds of the loci
dupGenotypes(msats, num.shared = 0.66)
```

The start and end positions and number of N's and indels can be generated with the _summarizeSeqs_ function:
```{r}
data(dolph.seqs)
seq.smry <- summarizeSeqs(as.DNAbin(dolph.seqs))
head(seq.smry)
```

Base frequencies can be generated with _baseFreqs_:
```{r}
bf <- baseFreqs(as.DNAbin(dolph.seqs))

# nucleotide frequencies by site
bf$site.freq[, 1:15]

# overall nucleotide frequencies
bf$base.freqs
```

Sequences can be scanned for low-frequency substitutions with _lowFreqSubs_:
```{r}
lowFreqSubs(as.DNAbin(dolph.seqs), min.freq = 2)
```

Unusual sequences can be identified by plotting likelihoods based on pairwise distances:
```{r}
data(dolph.haps)
haplotypeLikelihoods(as.DNAbin(dolph.haps))
```

All of the above functions can be conducted at once with the _qaqc_ function. Only those functions appropriate to the data type contained (haploid or diploid) will be run. Files are written for each output that are labelled either by the _\@description_ slot of the gtypes object or the optional _label_ argument of the function.